\chapter{Introduction}

%% SrJ: add explanations about profiling vs. imaging: size of beam and
%% SrJ: implications (ambiguities)

Underwater mapping and simulation are dual processes, while the latter produce
sonar responses for a given environment, the former use these responses to infer
the surroundings. As such simulation is a flexible way of generating data with a
known ground truth to test a mapping algorithm. However, to archive a correct
underwater simulation algorithm, simplifying assumptions on sound physics
and environmental characteristics are necessary, as well as a definition of the
sonar type being modeled.
%SLAM 

Profilings and imaging sonars are two classes of sonars whose differences lie
in the aperture of their sound beams. Profiling sonars have a narrow sound beam
and they are considered the laser scanner analog for underwater mapping,
even though profilings still have much wider beam than lasers. The simplest
approaches to underwater 3D mapping focus on applying laser scanner techniques
to profiling sonars, e.g. point cloud reconstruction.
% % SrJ: they don't really "give more information". They cover more space per %
% SrJ: beam, but the information is more ambiguous
 On the other hand, imaging sonars are usually
cheaper and have a wider sound beam, covering more space at the expense of
having a more ambiguous response. Thus, the choice of using imaging
sonars for mapping comes with the challenge of overcoming its measurement's
uncertainty.
%and benefiting from its extra information is a win-win strategy, the .
% % SrJ: "a win-win strategy" means that there are two parties ... I only see
% one

%Mapping is subject to different interpretations,


% % SrJ: apart from poor wording, the following is off the point. What you
% really
% % SrJ: want to stress is that there are various map representations. There are
% % SrJ: SLAMs on dense 3D maps / 2.5D ! Not every SLAM is feature-based, and a
% % SrJ: lot of feature-based SLAMs can generate dense maps as a byproduct
Besides stipulating a sonar type, the meaning of mapping ought to be narrowed
down. It is possible to generically define mapping as the process of gathering
multiple sensor data to characterize the surroundings. However, how this
characterization might be represented is dependent on the application.

A SLAM (\textbf{S}imultaneous \textbf{L}ocalization \textbf{a}nd
\textbf{M}apping) system has no intrinsic need for a human readable map.
In such a system, it could be interesting to store the map information only
through its most representative features, but even for SLAM that is not always
the case. It is often implemented as a grid with empty/full cells or even as a
continuous map.

The mapping of underwater environments is not just a part of a SLAM system. It
has importance on its own, it can be used for humans to visualize things that
could not be seeing otherwise. If the map is to be seen by a human it should
store and merge information about the environment, so that it can be displayed
as a usual map, 3D or 2D depending on the case. This representation also guides
how the data could be stored, e.g. if it wants to show a surface, it can be
stored as an elevation map, or if one wants to see a 3D object it can be stored
as a point cloud, a 3D grid, a continuous map, etc.

\input{Chap1/Motiv}
\input{Chap1/RevBib}
\input{Chap1/Obj}
\input{Chap1/Method}
\input{Chap1/WorkStruc}
