

\section{Methodology}
%\section{Metodology and Expected Results}
%But to get there it is important to describe some facts first.
 


% Using de aforementioned ideas, the main goal of the thesis reported here is to
% implement state of the art techniques for mapping, using binary bayesian
% filters. Fuse all the data on a octree structure using a robotic framework,
% named ROCK, and implement a visualization for the reconstruct underwater map.
% The data source shall be a imaging sonar mounted on a pan-tilt unit, so to
% provide the sonar extra degrees of freedom.
% 
% Optimization of the visible surface for based on the expected response of a
% sonar beam in a given direction will also be attempted. It might supply
% information about the surface material, specifically its reflectance. A
% technique based on previously articles.


The work is divided into tree main branches. Each having a theoretical research
(followed by implementation and testing ???).

The sonar model definition starts with a compilation on the description of sonar
variaties, sound waves physical properties on the water, reflection, sonar
directional gain and sources of noise. Then tests are realised on known
environments so to fit and validate the proposed model. The test results are
finally used create the full sonar model, including metric conventions by the
sonar protocol.


-------------------REVIEW-----------------------
The second branch deal with the map filling, basically the Binary Bayesian
Filter implementation. A review on Bayesian Filtering is scheduled before the
coding writing to be done on the robotics framework, ROCK. The implementation
will make use of the Octotree data structure, through the Octomap library, to
store the map.

The integration of the sonar model with the Bayesian Filter give the means to
process sonar data. So data acquired on the LNDC/UFRJ tank (Laboratório de
Ensaios Não Destrutivos, Corrosão e Soldagem), which loosely translated means
Non-Destructive Testing, Corrosion and Welding Laboratory, and on the Jirau
Power Plant, by means of the ROSA/COPPETEC project, will be processed and
compared to the tank and power plant entry layout, respectively.
 
In a more complex endeavor, which is the third branch, a theoretical derivation
for the optimization of 2D surfaces embed on 3D environments will take place.
It uses sonar response as expectation, and must take into account the
probabilistic nature of the map grid. The objective is, instead of having a
fixed sonar model, be able to better infer the underlying geometry of a sonar
responde, and also recover some information about the surrounding materials
reflectivity properties. This optimization algorithm will then be implemented on
ROCK so it can also be integrated to the Bayesian Filter, instead of the sonar
model. The same data processed by the combination ``Filter + Sonar Model'' shall
now be processed by ``Filter + Optimization''.
 
Based on the literature, even with not much similar works, the reconstruction
with \textit{a priori} sonar model will probably experience problems when
reconstructing corners, shallow angle surfaces, very complex (intense multipath)
or on highly noise environments. For the optimization, it shall encounter
similar issues, but a less accentuated shallow angle quality degradation and an
overall less blurry reconstruction.
---------------------------------------------------